---
title: "Calculating Walking trips impedance function- 1986"
---

# Calculating Walking trips impedance function- 1986

First load all of the libraries

```{r load-packages, include=FALSE, cache=FALSE}
#load all packages:
library(fitdistrplus) # Help to Fit of a Parametric Distribution to Non-Censored or Censored Data
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(here)
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(readxl) # Read Excel Files 
library(splitstackshape) # Stack and Reshape Datasets After Splitting Concatenated Values
library(tibble) # Simple Data Frames
library(readr) # Read Rectangular Text Data
library(tabulate) # Pretty Console Output for Tables
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
library(flextable) # Functions for Tabular Reporting
```

Here, the "walking_1986" dataset, which was constructed using the General Social Survey (GSS) dataset for walking trips in 1986, has been loaded.

```{r}
# Load the dataset
load(file = paste0(here(), "/data/walking_1986.Rda"))
```

Initially, we analyzed the travel time variable (DURATION) for 1986. The minimum recorded duration for walking trips is 1 minutes, while the maximum extends to 660 minutes. Furthermore, the average and median duration for walking trips are calculated to be 20.85 minutes and 10 minutes, respectively.

```{r}
summary(walking_1986$DURATION)
```

Subsequently, a weighted histogram was constructed. A histogram serves as a graphical representation that illustrates the frequency distribution of a dataset corresponding to a single variable. Typically, histograms organize data into 'bins' or 'range groups,' tallying the quantity of data points within each bin. This method effectively visualizes the distribution and variability of the dataset, providing insights into its underlying structure and patterns.

```{r}
ggplot(walking_1986, aes(x = DURATION )) + 
  geom_histogram(binwidth = 30) + 
  geom_vline(aes(xintercept=mean(DURATION)), color="blue", linetype="dashed", linewidth=1)
```

Here, we aim to analyze individuals who engaged in walking for more than 100 minutes, focusing on their origins and destinations. Additionally, we seek to ascertain the number of trips with a DURATION exceeding 100 minutes.

```{r}
walking_1986[walking_1986$DURATION > 100, ]
count_duration <- sum(walking_1986$DURATION > 100)
count_duration
```
As observed, out of 4,347 trips, 117 exceed 100 minutes in duration. Consequently, we have decided to exclude these trips from our analysis because they are all recreational.

```{r}
walking_1986 <- walking_1986 %>% filter(DURATION < 100)
summary(walking_1986$DURATION)
```


In our analysis to calculate the impedance function, we turned to the 'Fitdistrplus' package in R, an essential tool for fitting distributions to data. In the first step, the skewness-kurtosis plot was created. This plot is crucial for pinpointing the Pearson distribution that best matches our data. The process starts with the 'descdist' function, which calculates key descriptive statistics like the minimum, maximum, median, mean, standard deviation, skewness, and kurtosis.

Building on the groundwork laid by Cullen and Frey in 1999, our skewness-kurtosis plot helps us scrutinize the empirical distribution of our dataset. This plot clearly demarcates specific values or regions associated with common statistical distributions, aiding us in visually identifying the optimal model for our analysis. Distributions that exhibit fixed skewness and kurtosis values—such as normal, uniform, logistic, and exponential—appear as distinct points on the plot. This feature is particularly handy, as it indicates that data conforming to these distributions will align with those precise points, simplifying the identification process.

However, the plot also caters to distributions like gamma and lognormal, represented as lines, suggesting a variety of potential skewness and kurtosis values they can embody. This aspect is vital for understanding the flexibility these distributions offer when fitting data that may not align perfectly with a single model. Similarly, the beta distribution's broader areas on the plot signify its adaptability in form, accommodating a wider array of data shapes.

This visual tool, enriched by the comparative ease it brings to distribution selection, underscores the value of nuanced data analysis. By integrating these insights, we are better equipped to select the distribution that most accurately reflects the underlying characteristics of our data, enhancing the reliability and precision of our findings.

```{r creating Cullen and Frey graph for walkig 1986}
# creating a skew vs. kurtois graph for walking 1986

descdist(walking_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As illustrated in the Cullen and Frey graph, the gamma and exponential distributions appear to be the most suitable fits according to the aforementioned graph. Subsequently, we will conduct a comparative analysis of these models to ascertain the optimal model.

```{r}

# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution

gamma_w_1986_ <- fitdistrplus::fitdist(data=walking_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(walking_1986$WGHT_EPI))

fit_dgamma <- data.frame(f = dgamma(walking_1986$DURATION, shape = gamma_w_1986_$estimate[1], rate = gamma_w_1986_$estimate[2]), x = walking_1986$DURATION, type = "Gamma")

summary(gamma_w_1986_)

plot(x= fit_dgamma$x, y=fit_dgamma$f)


#Calculating weibull distribution
exp_w_1986_ <- fitdistrplus::fitdist(data=walking_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_1986$WGHT_EPI))

summary(exp_w_1986_)

fit_dexp <- data.frame(f = dexp(walking_1986$DURATION, rate = exp_w_1986_$estimate[1]), x = walking_1986$DURATION, type = "exp")

plot(x= fit_dexp$x, y=fit_dexp$f)

```

Now let's compare these models based on AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(walking_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(walking_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))

```

In the evaluation of different statistical models, we often rely on metrics such as the log-likelihood, Akaike Information Criterion (AIC), and Bayesian Information Criterion (BIC). The log-likelihood provides an estimate of the probability of observing our data under each model, with higher values being preferable. Both the AIC and BIC extend beyond mere goodness of fit, incorporating a penalty for excessive complexity to strike a balance between model accuracy and simplicity. The objective is to identify a model that not only fits the data well but also adheres to the principle of parsimony. Typically, models with lower AIC or BIC values are deemed more desirable.

Upon analysis of the gamma and exponential distribution outputs, it is observed that the gamma distribution achieves a superior log-likelihood and more favorable (lower) AIC and BIC scores. These findings indicate that the gamma distribution provides a closer fit to our data compared to the exponential distribution, and it does so without introducing undue complexity. In essence, the gamma distribution efficiently encapsulates the core characteristics of our data.


```{r}
walking_1986_gamma<- gamma_w_1986_
walking_1986 <- walking_1986 %>%
 mutate(f = dgamma(DURATION, walking_1986_gamma$estimate["shape"], walking_1986_gamma$estimate["rate"] )) 
summary(walking_1986$f)

```


### Destination : Home  (1)- 1986

The skewness vs. kurtosis graph provides descriptive statistics for walking trips to home destinations:

```{r creating Cullen and Frey graph for walkig 1986 - Home  (1)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_1_1986 <- walking_1986 %>%
  filter(destination == "1")   
descdist(w_1_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w1_1986_ <- fitdistrplus::fitdist(data=w_1_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_1_1986$WGHT_EPI))

fit_dgamma_w186 <- data.frame(f = dgamma(w_1_1986$DURATION, shape = gamma_w1_1986_$estimate[1], rate = gamma_w1_1986_$estimate[2]), x = w_1_1986$DURATION, type = "Gamma")

summary(gamma_w1_1986_)

plot(x= fit_dgamma_w186$x, y=fit_dgamma_w186$f)

#Calculating exponential distribution
exp_w1_1986_ <- fitdistrplus::fitdist(data=w_1_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_1_1986$WGHT_EPI))

summary(exp_w1_1986_)

fit_dexp_w186 <- data.frame(f = dexp(w_1_1986$DURATION, rate = exp_w1_1986_$estimate[1]), 
                       x = w_1_1986$DURATION, type = "exp")

plot(x = fit_dexp_w186$x, y = fit_dexp_w186$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_1_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_1_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_186_gamma <- gamma_w1_1986_
w_1_1986 <- w_1_1986 %>%
 mutate(f_1 = dgamma(DURATION, walking_186_gamma$estimate["shape"], walking_186_gamma$estimate["rate"])) 
summary(w_1_1986$f_1)
```


### Destination : Work place (2)- 1986

The skewness vs. kurtosis graph provides descriptive statistics for walking trips to work and school destinations:

```{r creating Cullen and Frey graph for walkig 1986 - Work place (2)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_2_1986 <- walking_1986 %>%
  filter(destination == "2")   
descdist(w_2_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w2_1986_ <- fitdistrplus::fitdist(data=w_2_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_2_1986$WGHT_EPI))

fit_dgamma_w286 <- data.frame(f = dgamma(w_2_1986$DURATION, shape = gamma_w2_1986_$estimate[1], rate = gamma_w2_1986_$estimate[2]), x = w_2_1986$DURATION, type = "Gamma")

summary(gamma_w2_1986_)

plot(x= fit_dgamma_w286$x, y=fit_dgamma_w286$f)

#Calculating exponential distribution
exp_w2_1986_ <- fitdistrplus::fitdist(data=w_2_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_2_1986$WGHT_EPI))

summary(exp_w2_1986_)

fit_dexp_w286 <- data.frame(f = dexp(w_2_1986$DURATION, rate = exp_w2_1986_$estimate[1]), 
                       x = w_2_1986$DURATION, type = "exp")

plot(x = fit_dexp_w286$x, y = fit_dexp_w286$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_2_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_2_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_286_gamma <- gamma_w2_1986_
w_2_1986 <- w_2_1986 %>%
 mutate(f_2 = dgamma(DURATION, walking_286_gamma$estimate["shape"], walking_286_gamma$estimate["rate"])) 
summary(w_2_1986$f_2)
```

### Destination : Other's Home (3)- 1998

The skewness vs. kurtosis graph provides descriptive statistics for walking trips to other's home destinations:

```{r creating Cullen and Frey graph for walkig 1986 - Other Home (3)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_3_1986 <- walking_1986 %>%
  filter(destination == "3")   
descdist(w_3_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w3_1986_ <- fitdistrplus::fitdist(data=w_3_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_3_1986$WGHT_EPI))

fit_dgamma_w386 <- data.frame(f = dgamma(w_3_1986$DURATION, shape = gamma_w3_1986_$estimate[1], rate = gamma_w3_1986_$estimate[2]), x = w_3_1986$DURATION, type = "Gamma")

summary(gamma_w3_1986_)

plot(x= fit_dgamma_w386$x, y=fit_dgamma_w386$f)



#Calculating exponential distribution
exp_w3_1986_ <- fitdistrplus::fitdist(data=w_3_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_3_1986$WGHT_EPI))

summary(exp_w3_1986_)

fit_dexp_w386 <- data.frame(f = dexp(w_3_1986$DURATION, rate = exp_w3_1986_$estimate[1]), 
                       x = w_3_1986$DURATION, type = "exp")

plot(x = fit_dexp_w386$x, y = fit_dexp_w386$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_3_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_3_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_386_gamma <- gamma_w3_1986_
w_3_1986 <- w_3_1986 %>%
 mutate(f_3 = dgamma(DURATION, walking_386_gamma$estimate["shape"], walking_386_gamma$estimate["rate"])) 
summary(w_3_1986$f_3)
```



In the final step, we constructed a new walking data frame, "walking_1986_f", that contains all the 'f' values for walking trips in 1986.

```{r}
walking_1986_f <- bind_rows(w_1_1986, w_2_1986, w_3_1986)

```

Save:
```{r}
save(walking_1986_f,
     file = paste0(here(), "/data/walking_1986_f.Rda"),
     compress = "bzip2",
     version = 2,
     ascii = FALSE)
```



