---
title: "Calculating Walking trips impedance function- 1986"
---

# Active-transport walking and cycling behaviour in Canada

## Introduction

## Data and Method

Data on walking and cycling trips to different destinations was derived from Canada's General Social Survey (GSS) from 1986 to 2015. GSS is a social trends survey in order to monitor changes in the living conditions and well-being of Canadians and to provide information on specific social policy issues. one of the important information of GSS is about Time-use surveys. This survey collects information on all human activities and can therefore inform a broad range of policies. Statistics Canada has been conducting time-use surveys since 1986 at approximately five- to seven- year intervals, most recently in 2015.The GSS on time use employs a retrospective 24-hour time diary to collect information on respondents' participation in, and time spent on, a wide variety of day-to-day activities. In addition, information is collected on the location where these activities occurred (e.g., at home, at work, etc.).in addition, This dataset contains travel time data of Many of the Census Metropolitan Areas (CMAs) and non-CMA areas all over Canada. CMAs are including St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver. and the non-CMA areas of each of the ten provinces were also grouped to form ten more strata.

walking and cycling data are categorized based on the three different destinations such as home, school and work, and others' homes. every episode includes, the start and end time of each event, duration, weight, and concurrent contextual information, such as the location of the activities (e.g., home, work, schools,...).

complete method

# Result

First load all of the libraries

```{r load-packages, include=FALSE, cache=FALSE}
#load all packages:
library(fitdistrplus) # Help to Fit of a Parametric Distribution to Non-Censored or Censored Data
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(readxl) # Read Excel Files 
library(splitstackshape) # Stack and Reshape Datasets After Splitting Concatenated Values
library(tibble) # Simple Data Frames
library(readr) # Read Rectangular Text Data
library(tabulate) # Pretty Console Output for Tables
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
library(flextable) # Functions for Tabular Reporting
```

Here, we are loading all the primary files of the GSS dataset to create our database for analysis.

```{r}
# Load the dataset
load("data/walking_1986.Rda")
#list.files("data")
```


### Calculating impedance function for walking in 1986

```{r}

count <- sum(walking_1986$DURATION > 100)
print (count)

walking_1986 <- walking_1986 %>% filter(DURATION < 100)
summary(walking_1986$DURATION)
```

using a skew vs. kurtois graph to descriptive statistics of data:

```{r creating Cullen and Frey graph for walkig 1986}
# creating a skew vs. kurtois graph for walking 1986

descdist(walking_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```
As shown in the Cullen and Frey graph, we see that Weibull, gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.


```{r}

# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution

gamma_w_1986_ <- fitdistrplus::fitdist(data=walking_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(walking_1986$WGHT_EPI))

fit_dgamma <- data.frame(f = dgamma(walking_1986$DURATION, shape = gamma_w_1986_$estimate[1], rate = gamma_w_1986_$estimate[2]), x = walking_1986$DURATION, type = "Gamma")

summary(gamma_w_1986_)

plot(x= fit_dgamma$x, y=fit_dgamma$f)


#Calculating weibull distribution
exp_w_1986_ <- fitdistrplus::fitdist(data=walking_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_1986$WGHT_EPI))

summary(exp_w_1986_)

fit_dexp <- data.frame(f = dexp(walking_1986$DURATION, rate = exp_w_1986_$estimate[1]), x = walking_1986$DURATION, type = "exp")

plot(x= fit_dexp$x, y=fit_dexp$f)

```


Now let's compare these models based on AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(walking_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(walking_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))

```

```{r}
walking_1986_gamma<- gamma_w_1986_
walking_1986 <- walking_1986 %>%
 mutate(f = dgamma(DURATION, walking_1986_gamma$estimate["shape"], walking_1986_gamma$estimate["rate"] )) 
summary(walking_1986$f)

```


### Destination : Home  (1)- 1986

using a skew vs. kurtois graph to descriptive statistics of data

```{r creating Cullen and Frey graph for walkig 1986 - Home  (1)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_1_1986 <- walking_1986 %>%
  filter(destination == "1")   
descdist(w_1_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w1_1986_ <- fitdistrplus::fitdist(data=w_1_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_1_1986$WGHT_EPI))

fit_dgamma_w186 <- data.frame(f = dgamma(w_1_1986$DURATION, shape = gamma_w1_1986_$estimate[1], rate = gamma_w1_1986_$estimate[2]), x = w_1_1986$DURATION, type = "Gamma")

summary(gamma_w1_1986_)

plot(x= fit_dgamma_w186$x, y=fit_dgamma_w186$f)

#Calculating exponential distribution
exp_w1_1986_ <- fitdistrplus::fitdist(data=w_1_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_1_1986$WGHT_EPI))

summary(exp_w1_1986_)

fit_dexp_w186 <- data.frame(f = dexp(w_1_1986$DURATION, rate = exp_w1_1986_$estimate[1]), 
                       x = w_1_1986$DURATION, type = "exp")

plot(x = fit_dexp_w186$x, y = fit_dexp_w186$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_1_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_1_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_186_gamma <- gamma_w1_1986_
w_1_1986 <- w_1_1986 %>%
 mutate(f_1 = dgamma(DURATION, walking_186_gamma$estimate["shape"], walking_186_gamma$estimate["rate"])) 
summary(w_1_1986$f_1)
```


### Destination : Work place (2)- 1986

using a skew vs. kurtois graph to descriptive statistics of data

```{r creating Cullen and Frey graph for walkig 1986 - Work place (2)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_2_1986 <- walking_1986 %>%
  filter(destination == "2")   
descdist(w_2_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w2_1986_ <- fitdistrplus::fitdist(data=w_2_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_2_1986$WGHT_EPI))

fit_dgamma_w286 <- data.frame(f = dgamma(w_2_1986$DURATION, shape = gamma_w2_1986_$estimate[1], rate = gamma_w2_1986_$estimate[2]), x = w_2_1986$DURATION, type = "Gamma")

summary(gamma_w2_1986_)

plot(x= fit_dgamma_w286$x, y=fit_dgamma_w286$f)

#Calculating exponential distribution
exp_w2_1986_ <- fitdistrplus::fitdist(data=w_2_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_2_1986$WGHT_EPI))

summary(exp_w2_1986_)

fit_dexp_w286 <- data.frame(f = dexp(w_2_1986$DURATION, rate = exp_w2_1986_$estimate[1]), 
                       x = w_2_1986$DURATION, type = "exp")

plot(x = fit_dexp_w286$x, y = fit_dexp_w286$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_2_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_2_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_286_gamma <- gamma_w2_1986_
w_2_1986 <- w_2_1986 %>%
 mutate(f_2 = dgamma(DURATION, walking_286_gamma$estimate["shape"], walking_286_gamma$estimate["rate"])) 
summary(w_2_1986$f_2)
```

### Destination : Other's Home (3)- 1998

using a skew vs. kurtois graph to descriptive statistics of data

```{r creating Cullen and Frey graph for walkig 1986 - Other Home (3)}
# creating a skew vs. kurtois graph for walking 1986

# Filter the data based on the "destination" column
 w_3_1986 <- walking_1986 %>%
  filter(destination == "3")   
descdist(w_3_1986$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

As shown in the Cullen and Frey graph, we see that gamma, weibull and exponential will likely be the best fit according to the graph above. Then we compare these models to find the best model.

```{r}
# Based on the skew vs. kurtois graph the best distribution

#Calculating gamma distribution 
gamma_w3_1986_ <- fitdistrplus::fitdist(data=w_3_1986$DURATION%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead", weights = round(w_3_1986$WGHT_EPI))

fit_dgamma_w386 <- data.frame(f = dgamma(w_3_1986$DURATION, shape = gamma_w3_1986_$estimate[1], rate = gamma_w3_1986_$estimate[2]), x = w_3_1986$DURATION, type = "Gamma")

summary(gamma_w3_1986_)

plot(x= fit_dgamma_w386$x, y=fit_dgamma_w386$f)



#Calculating exponential distribution
exp_w3_1986_ <- fitdistrplus::fitdist(data=w_3_1986$DURATION%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(w_3_1986$WGHT_EPI))

summary(exp_w3_1986_)

fit_dexp_w386 <- data.frame(f = dexp(w_3_1986$DURATION, rate = exp_w3_1986_$estimate[1]), 
                       x = w_3_1986$DURATION, type = "exp")

plot(x = fit_dexp_w386$x, y = fit_dexp_w386$f)

```

Now let's compare these models AIC and BIC to see which fits the walk data best:

```{r}
#comparing these models AIC and BIC to see which fits the walk data best:
broom::glance(MASS::fitdistr(w_3_1986$DURATION%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(w_3_1986$DURATION%>% unlist() %>% as.numeric(),"exponential"))
```

So, gamma has largest logLik and the smallest AIC and BIC. we will pick gamma function for walking trips in 1986!

```{r}
walking_386_gamma <- gamma_w3_1986_
w_3_1986 <- w_3_1986 %>%
 mutate(f_3 = dgamma(DURATION, walking_386_gamma$estimate["shape"], walking_386_gamma$estimate["rate"])) 
summary(w_3_1986$f_3)
```



creating a new walking data frame that have all the f values for walking 1986

```{r}
walking_1986_f <- bind_rows(w_1_1986, w_2_1986, w_3_1986)

```

Save:
```{r}
save(walking_1986_f,
     file = paste0(here(), "/data/walking_1986_f.rda"),
     compress = "bzip2",
     version = 2,
     ascii = FALSE)
```



